\documentclass[../main.tex]{subfiles}

\begin{document}
\section{Conclusions}\label{sec:conclusions}
This article has demonstrated that locally injected turbulence in a 2D space spreads to the boundaries of the domain, provided the Reynolds number is sufficiently high. This is in contrast to previous works in 3D domains, where the energy is shown to be confined to the region of the forcing. The results of the Navier-Stokes simulations show that the enstrophy density distribution is consistent with the point vortex simulations, where the vortices spreads across the domain and reaches the boundaries. This is observed for different values of the size of the forcing region $k_r$ and the Reynolds number. Moreover, the profile of the energy and enstrophy density distributions as a function of the radius from the center of the domain has been claimed to follow a power law of $A/r^2$, for some constant $A$. Particularly in the case of enstrophy, the results show remarkable consistency, suggesting that as $k_r\to 0$ and sufficient time elapses, the power law for the enstrophy density distribution is expected to cover the entire range of $r$.

Regarding the point vortex simulations, the findings indicate that the power law $A/r$ provides a good approximation for the density distribution of vortices per unit length $\rho_r$ This contrasts with the results from Navier-Stokes simulations, where the enstrophy density distribution is better described by a power law $A/r^2$. This discrepancy implies, as illustrated in the figures above, a faster decay of enstrophy in Navier-Stokes simulations compared to the point-vortex model. This suggests the need to identify a more accurate metric for comparing both models.

For an extension of this work, it would be interesting to run the simulations for a longer time and higher values of $k_r$ and $\Re$ for the embarrassingly parallel code. This would allow for a more accurate determination of the accordance of the plots for the evolution of the mean radii in both the Navier-Stokes and the point vortex simulations.
Finally, these results could be replicated with the addition of a drag force $-\alpha \vf{u}$ to the Navier-Stokes equations, where $\alpha$ is a constant. This would allow us to investigate whether the energy still spreads as far as the domain allows, even in the presence of the drag force. To account that in the point vortex model, an extra equation is suggested to be added to the pair of equations for $x_i$ and $y_i$. This equation would concentrate the drag effect on the intensity of circulation of the point vortex, formulated as
\begin{equation}
	\dot{\Gamma}_i = -\beta \Gamma_i
\end{equation}
for some constant $\beta$.
\end{document}
